### 2015-03-14 intro.tex
various updates on abstract and intro

Linux kernel modules (e.g. device drivers and filesystems) exist at the thin boundary between hardware and software. They are complex pieces of system-level software responsible for the interaction between a user application, the operating system and any hardware devices that are attached to a computer. Kernel modules are notoriously hard to develop and even harder to debug~\cite{corbet2005linux}. Especially in the case of drivers, this has a negative impact in hardware product releases, as time-to-market is commonly dominated by device driver development, verification and validation schedules~\cite{yavatkar2012era}.

@@

In this paper, we present \emph{static pairwise lockset analysis}, a novel technique for automatically verifying the absence of data races in Linux kernel modules (e.g. device drivers and filesystems). The key idea behind our approach is that a module can be proven free from data races by (i) deriving a sound \emph{sequential} model that \emph{over-approximates} the originally concurrent program, (ii) instrumenting it for lockset analysis and race checking, and (iii) asserting that all accesses to the same shared resource are protected by at least one common lock. The immediate benefit is that our approach not only avoids reasoning about thread interleavings, and thus has the potential to scale well, but also allows the reuse of existing successful sequential verification techniques.

We then show that by exploiting the race-freedom guarantees that our static analysis provides, we can achieve a sound partial-order reduction and significantly accelerate \corral~\cite{lal2012solver}, a state-of-the-art bug finder that is used by Microsoft to analyse device drivers. We have prototyped this approach in \whoop, a practical tool for automatic concurrency verification, which we used to analyse 10 modules available in the Linux distribution.

To summarize, our contributions are as follows:
\begin{itemize}
\item We present static pairwise lockset analysis,  a sound scalable technique for automatically verifying absence of data races in Linux kernel modules.
\item We show that we can achieve a sound partial-order reduction using our technique and significantly accelerate \corral, a state-of-the-art bug finder.
\item We develop \whoop, a prototype tool that implements our approach and we use it to analyse 10 Linux kernel modules.
\end{itemize}

>>>>

Linux kernel modules, such as device drivers and filesystems, exist at the thin boundary between hardware and software. They are complex pieces of system-level software responsible for the interaction between a user application, the operating system and any hardware devices that are attached to a computer. Linux kernel modules are notoriously hard to develop and even harder to debug~\cite{corbet2005linux}. Especially in the case of drivers, this has a negative impact in hardware product releases, as time-to-market is commonly dominated by driver development, verification and validation schedules~\cite{yavatkar2012era}.

@@

In this paper, we present \whoop, a novel infrastructure for (i) automatically analysing the concurrent behaviour of Linux kernel modules written in the C language and (ii) accelerating external bug finders that can easily plug in our toolchain. \whoop is empowered by \emph{static pairwise lockset analysis}, a novel technique for automatically verifying the absence of data races in Linux kernel modules (e.g. device drivers and filesystems). The key idea behind our approach is that a module can be proven free from data races by (i) deriving a sound \emph{sequential} model that \emph{over-approximates} the originally concurrent program, (ii) instrumenting it for lockset analysis and race checking, and (iii) asserting that all accesses to the same shared resource are protected by at least one common lock. The immediate benefit is that our approach not only avoids reasoning about thread interleavings, and thus has the potential to scale well, but also allows the reuse of existing successful sequential verification techniques.

We then show that by exploiting the race-freedom guarantees that our static analysis provides, we can achieve a sound partial-order reduction that significantly accelerates \corral~\cite{lal2012solver}, a state-of-the-art bug finder that is used by Microsoft to analyse device drivers and other concurrent programs. Using \whoop and \corral we analysed 10 modules available in the Linux distribution.

To summarize, our contributions are as follows:
\begin{itemize}
\item We present \whoop, a practical tool for analysing the concurrent behaviour of Linux kernel modules and accelerating a plugged-in precise bug finder.
\item We discuss static pairwise lockset analysis,  a sound scalable technique for automatically verifying absence of data races in Linux kernel modules.
\item We show that we can achieve a sound partial-order reduction using our technique and significantly accelerate \corral, a state-of-the-art bug finder.
\item We analyse 10 Linux kernel modules, including various drivers and a filesystem.
\end{itemize}

### 2015-03-14 theory.tex
various updates on abstract and intro

The proof sketch behind our static lockset analysis is given in Section~\ref{proof}. The verification technique is described in more detail in Section~\ref{method}. The prototype implementation of our approach as the practical tool \textsc{Whoop} is presented in Section~\ref{implementation}.

\subsection{Static lockset analysis proof sketch}
\label{proof}

@@

\subsection{Verification method}
\label{method}

@@

Our approach begins by performing \emph{two-thread reduction}, an abstraction that removes all but two arbitrary threads, each running an entry point of the originally concurrent driver, and then performs \emph{pairwise sequentialisation}, which combines the two arbitrary threads in a single sequential pair. This is achieved by replacing the \texttt{return} statement of the first entry point with a \texttt{goto} statement that jumps execution to the first statement of the second entry point. This process repeats until all pairs of entry points have been sequentialised. To achieve soundness, each time an entry point performs a read access to a shared resource, we return a nondeterministic value. This over-approximates any effects from all the unmodeled threads on the device driver shared state.

@@

>>>>

The proof sketch behind our static lockset analysis is given in Section~\ref{proof}. The verification technique is described in more detail in Section~\ref{method}. The summarisation that we perform to achieve scalability is discussed in Section~\ref{summarisation}. Various optimisations that we perform to achieve precision are described in Section~\ref{optimisations}. Finally, the prototype implementation of our approach as the practical tool \textsc{Whoop} is presented in Section~\ref{implementation}.

\subsection{Static Lockset Analysis Proof Sketch}
\label{proof}

@@

\subsection{Verification Method}
\label{method}

@@

Our verification technique begins by performing \emph{two-thread reduction}, a sound abstraction that removes all but two arbitrary threads, each running an entry point of the originally concurrent program, and then performs \emph{pairwise sequentialisation}, which combines the two arbitrary threads in a single sequential pair. This is achieved by replacing the \texttt{return} statement of the first entry point with a \texttt{goto} statement that jumps execution to the first statement of the second entry point. This process repeats until all pairs of entry points have been sequentialised. To achieve soundness, each time an entry point performs a read access to a shared resource, we return a nondeterministic value. This over-approximates any effects from all the unmodeled threads on the device driver shared state.

@@

\subsection{Watchdog Summarisation}
\label{summarisation}

Draft.

\subsection{Optimisations}
\label{optimisations}

Draft.
